---
title: "Analysis_IMDB_MovieDataset"
author: "Sam Kazan"
date: "2023-01-20"
output:
  html_document:
    df_print: paged
---

```{r setup,}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, comment = NA)
```

```{r result = FALSE}
library(ggplot2movies)
library(dplyr)
library(ggplot2)
data(movies)
##movies
```

# Questions to be Adressed 

1. What is the range of years of production of the movies of this data set (i.e. what is the year of production of the oldest movie and of the most recent movie in this data set)?
2. What proportion of movies have their budget included in this data base, and what proportion doesn’t? What are top 5 most expensive movies in this data set?
3. What are top 5 longest movies?
4. Of all short movies, which one is the shortest (in minutes)? Which one is the longest? How long are the shortest and the longest short movies?
 library(ggplot2movies) data(movies)
5. How many movies of each genre (action, animation, comedy, drama, documentary, romance, short) are there in this data base? (use a bar plot)
6. What is the average rating of all movies within each genre? (use a bar plot)
7. What is the average rating of all movies within each genre that were produced in the years 2000-2005?
8. For each of the first 6 genres (not including short movies) consider only movies from 1990 until the last year recorded and plot a function of the number of movies in this data base of corresponding genre produced by year, for years from 1990 until the last year recorded. For each of the 6 genres you should have one curve, and plot all the curves in the same figure. Naturally, use different colors, and appropriate legend.
9. Finally, formulate 3 questions of your choice.

## Question 1

```{r}
range_years <- range(movies %>% select(year) %>% pull())
```


Movies in the database range from the year `r range_years[1]` to year `r range_years[2]`

## Question 2

```{r}
budget_NA <- movies %>% filter(is.na(budget)) %>% nrow()
budget_no_NA <- movies %>% filter(!is.na(budget)) %>% nrow()
budgetcount <- c(budget_NA,budget_no_NA)
total_movie <- movies %>%  nrow()

```

There are `r budget_NA` movies that does not have a defined total budget while `r budget_no_NA` movies have listed total budget number. `r 100*budget_no_NA/total_movie` % of the movies don't have a budget included in the dataset. On the other hand `r 100*budget_NA/total_movie` % of the movies have a budget included.

```{r}
dftop5 <- movies %>% 
filter(!is.na(budget)) %>% 
arrange(desc(budget)) %>% 
  top_n(5, budget) %>% 
  select(title, budget)

ggplot(data=dftop5, aes(x = title, y = budget)) + 
  geom_bar(stat = "identity", fill = "grey") + 
  xlab("Title") + ylab("Budget") + ggtitle("Top 5 Movies by Budget") +
  theme_classic() +
  coord_flip() 

TopMoviesByBudget <- dftop5 %>% pull(title)
```



Top 5 most expensive movies in this data set are `r TopMoviesByBudget`


## Question 3 & 4

```{r}
dflongest <-  movies %>% 
              select(title, length) %>%
              arrange(desc(length)) %>%
              top_n(5, length) 
LongestMovies <- dflongest %>% pull(title)

dfshorts <- movies %>% 
            select(title, length, Short) %>%
            filter(Short == 1) %>% 
            arrange(desc(length))


shortmovies <- dfshorts %>%
  select(length) %>%
  filter(length == 1) %>%
  summarize(count = n()) 

ggplot(data=dflongest, aes(x = title, y = length)) + 
  geom_bar(stat = "identity", fill = "magenta") + 
  xlab("Title") + ylab("Length") + ggtitle("Top 5 Movies by Length") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

```

Top 5 longest movies are `r LongestMovies`. The longest movies are `r LongestMovies[1]` and `r LongestMovies[3]` which triple the length of the other 3 movies that are in the top 5 list.\newline

Looking at all the short movies at the list the longest short movie is `r dfshorts %>% pull(title) %>% first()` with the running time `r dfshorts %>% pull(length) %>% first()` mins. There are `r shortmovies` short movies that have the running time of `r dfshorts %>% pull(length) %>% last()` mins. Those movies are the shortest short movies in the database. 

## Question 5

```{r}

dfGenres <- movies %>%
  mutate(genres = case_when(
    Action == 1 ~ "action",
    Animation == 1 ~ "animation",
    Comedy == 1 ~ "comedy",
    Drama == 1 ~ "drama",
    Documentary == 1 ~ "documentary",
    Romance == 1 ~ "romance",
    Short == 1 ~ "short"
  )) %>% 
  filter(!is.na(genres))

dfGenres %>%
  group_by(genres) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = genres, y = count)) + 
  ggtitle("Distribution of movies across genres.")+
  geom_col(fill = "tomato3") +
  geom_text(aes(label = count), position = position_dodge(),vjust = -0.25, color = "black")
  
genre_coutn <- dfGenres %>%
  group_by(genres) %>%
  count() %>%
  pull()

genres_list = dfGenres %>% distinct(genres) %>% arrange(genres) %>% pull(genres)  

```



There are 7 classification of movies: `r genres_list `. NA are the ones without a specific genre listed in the database and are not included int the plot. Comedy and Drama are the most produced genres with movie counts of `r genre_coutn[3]` and `r genre_coutn[5]`, respectively.   

## Question 6 

```{r}
dfGenres_avg <- dfGenres %>% 
                group_by(genres) %>% 
                summarize(avg_rating = round(mean(rating), 2))
##dfGenres_avg
average_rating = dfGenres_avg %>% pull(avg_rating)
 
ggplot(data = dfGenres_avg, aes(x = genres, y = avg_rating)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Ratings by Genre")+
  xlab("Genre") +
  ylab("Average Rating") +
  geom_col(fill = "lightblue") +
  geom_text(aes(label = avg_rating), position = position_fill(vjust = 2), color = "black") +
  scale_linetype_manual(name = "Average", values = c("Average" = "dashed"))
  
```

All of the genres have average ratings between `r range(dfGenres_avg$avg_rating)[1]` and `r range(dfGenres_avg$avg_rating)[2]`. Documentaries and Dramas are most liked movie genres with average ratings `r average_rating[4]` and `r average_rating[5]` consecutively.

## Question 7

```{r}
df_genres2000 <- dfGenres %>% 
  filter( year >+ 2000, year<2006) %>%
  group_by(genres) %>% 
  summarize(avg_rating_2000 = round(mean(rating),2))


ggplot(data=df_genres2000, aes(x = genres, y = avg_rating_2000)) +
  geom_col(fill= "lightgreen") +
  geom_text(aes(label = avg_rating_2000 ), position = position_fill(, vjust = 2.5)) +
  xlab("Genre") +
  ylab("Average Rating") +
  ggtitle("Average Rating by Genre for movies produced in 2000-2005")
```
Looking at the movies produced between 2000-2005 on average the genres that are most liked were `r genres_list[4]` and `r genres_list[7]` movies. Comparing this graph with all the averages by genres, we can see that `r genres_list[7]` movies started getting ranked higher which surpass `r genres_list[5]` from previous dataset. 

## Question 8

```{r}
movies_1990_onwards <- dfGenres %>% 
                        filter(year >= 1990,year < 2005,genres != "NA", 
                               genres != "short") %>% 
                        group_by(genres, year) %>% 
                        count()
##dfGenres
##movies_1990_onwards

ggplot(data = movies_1990_onwards, aes(x = year, y = n, color = genres)) +
  geom_line() +
  scale_color_manual(values = c("red", "blue", "green", "purple", 
                                     "orange", "black")) +
  labs(title = "Number of movies by genre and year", x = "Year",
       y = "Number of movies", color = "Genre")
```
The graph shows that the number of movies released in genres `r genres_list[1]` and `r genres_list[3]`are always higher (at least double) than the movies rleased under other genres.

# Question 9

## Average Budget by Genres

Question : Which genres have the biggest budget comparing all 7 genres ?

Calculations are for budget >= 1000000
```{r}
dfBudget <- dfGenres %>% 
  group_by(genres) %>% 
  filter(budget != "NA", budget <= 1000000) %>% 
  mutate(budget_in_mil = budget/1000000) %>% 
  summarize(avg_budget = mean(budget)) %>% 
  invisible()

budgetPerM = dfBudget %>% pull(avg_budget) 

  ggplot(data = dfGenres, aes(x=genres,y=budget))+ 
  geom_boxplot(fill="purple") + 
  labs(y='Budget by Millions',title='Average Budget by Genre')  
  
```
The graphs shows that action and animation genres have greater spending than all other genres.


## Average votes by rating

Question : Is there a connection between number of votes and the rating of a movie?


```{r}
dfvotes <- dfGenres %>% 
  filter(!is.na(votes)) %>%  
  group_by(rating) %>% 
  summarize(mean_votes = mean(votes))

ggplot(data = dfvotes, aes(x = rating, y = mean_votes)) +
  geom_bar(stat = 'identity') +
  geom_smooth(method = "loess", se = F) +
  labs(x = 'rating', y='votes',title='Average votes by rating')
```
Looking at the distribution of votes among the rating we can observe  the most voted movies are rated between 8.0 and 9.0. The number of movies that received more than 9.0 rating are scarce.

## Top 5 voted movies by their rating values

Question : What percent of the ratings are given for the most voted movies ?

```{r}
dfVoters <- dfGenres %>% 
  arrange(desc(votes)) %>% 
  top_n(5,votes)

dfr_main <- dfVoters
dfr_temp <- dfr_main %>% 
  mutate(r_value = r1) %>% 
  mutate(r_label = 1)
dfr_main_1 <- bind_rows(dfr_main, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r2) %>% 
  mutate(r_label = 2)
dfr_main_2 <- bind_rows(dfr_main_1, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r3) %>% 
  mutate(r_label = 3)
dfr_main_3 <- bind_rows(dfr_main_2, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r4) %>% 
  mutate(r_label = 4)
dfr_main_4 <- bind_rows(dfr_main_3, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r5) %>% 
  mutate(r_label = 5)
dfr_main_5 <- bind_rows(dfr_main_4, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r6) %>% 
  mutate(r_label = 6)
dfr_main_6 <- bind_rows(dfr_main_5, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r7) %>% 
  mutate(r_label = 7)
dfr_main_7 <- bind_rows(dfr_main_6, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r8) %>% 
  mutate(r_label = 8)
dfr_main_8 <- bind_rows(dfr_main_7, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r9) %>% 
  mutate(r_label = 9)
dfr_main_9 <- bind_rows(dfr_main_8, dfr_temp)

dfr_temp <- dfr_main %>% 
  mutate(r_value = r10) %>% 
  mutate(r_label = 10)
dfr_main_10 <- bind_rows(dfr_main_9, dfr_temp)

dfr_main_10 <- dfr_main_10 %>% 
  filter(!is.na(r_value))
##dfr_main_10

ggplot(data = dfr_main_10, aes(x=title, y = r_value, fill = r_label)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y='% Rating',title='Top 5 movies by R values', x='Title')+ 
  scale_fill_gradient(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))+
  coord_flip() 
 
```
This plot shows that most people voted the same for the top 5 movies considering the first 8 percentile ratings. Interestingly except Lord of The Rings people voted more nines over tens.Lord of The Rings and Shawshank Redemption received most amount of tens among the top five voted movies.
